\newcommand{\ro}{\raisebox{0.25\baselineskip}{\tiny{\textregistered{}}}}

\begin{thebibliography}{99}
\bibitem{Abraham08}Erika \'Abrah\'am, Immo Grabe, Andreas Gr\"uner, Martin Steffen,
\textit{Behavioral interface description of an object-oriented language with futures and promises},
Elsevier,
2008.

\bibitem{Alhadidi07}Dima Alhadidi, Nadia Belblidia, Mourad Debbabi, Prabir Bhattacharya,
\textit{$\lambda$\_AOP: An AOP Extended Lambda-Calculus},
CSL-CIISE, Concordia University,
2007.

\bibitem{Dabek02}Frank Dabek, Nickolai Zeldovich, Frans Kaashoek, David Mazières, Robert Morris,
\textit{Event-driven Programming for Robust Software},
MIT Laboratory for Computer Science,
2002.

\bibitem{Goetz10}Brian Goetz, Tim Peierls, Joshua Bloch, Joseph Bowbeer, David Holmes, Doug Lea,
\textit{Java Concurrency in Practice},
Addison-Wesley,
2010.

\bibitem{Gosling14}James Gosling, Bill Joy, Guy Steele, Gilad Bracha, Alex Buckley,
\textit{The Java\ro Language Specification (Java SE 8 Edition)},
Oracle America,
2014.

\bibitem{Lauer78}Lauer H.C., Needham R.M.,
\textit{On the Duality of Operating Systems Structures},
inProc. Second International Symposium on Operating Systems,
1978.

\bibitem{Lindholm14}Tim Lindholm, Frank Yellin, Gilad Bracha, Alex Buckley,
\textit{The Java\ro Virtual Machine Specification (Java SE 8 Edition)},
Oracle America,
2014.

\bibitem{Maier12}Ingo Maier, Martin Odersky,
\textit{Deprecating the Observer Pattern with Scala.React},
LAMP-EPFL,
2012.

\bibitem{Niehren06}Joachim Niehren, Jan Schwinghammer, Gert Smolka,
\textit{A Concurrent Lambda Calculus with Futures},
Theoretical Computer Science, Elsevier,
2006.

\bibitem{Wang09}Meng Wang, Bruno C. d. S. Oliveira,
\textit{What Does Aspect-Oriented Programming Mean for Functional Programmers},
Oxford University Computing Laboratory,
2009.
\end{thebibliography}
