
\chapter{Zone your Java}
\label{ch:impl}

This chapter closes the Zone presentation with a brief overview of its implementation and usage.
The aim is to give you an insight on the Zone architecture and get you familiarized with some utilization tips and tricks.

\section{Idioms}

Here is a small digressions about common used code patterns that may trouble a regular Java developer using the Zone.

\begin{itemize}
\item $\lambda$s typing and creation (especially for hooks)
\item Key-typed-map and Heap pollution concerns
\item The match function (use Option class for Zone values lookup)
\end{itemize}

\section{Make your Zone}

\begin{itemize}
\item Zone class extension - implement your hooks
  \begin{itemize}
  \item cross in - out method
  \item get around hooks (sync and async) method
  \item get value method
  \end{itemize}
\item Use embedded zone (value zone, around zone, b-r-a zone, cross zone, error zone)
\item Build your own zone property (as in last point of application)
\end{itemize}


\section{Zoned Executors}

As evoked in (some previous chapter), the Zone integrates on top of an asynchronous execution framework. In order to make zone fully functional (regarding asynchronous invocations), you will need either

Use an instrumented execution framework :
\begin{itemize}
\item Executor, ExecutorService, CompletableFuture (with limitations of replacing JRE classes by others : different packages/naming
\item instrumented third part library, \vertx\ example.
\end{itemize}

Or instrument your own execution framework :
\begin{itemize}
\item Think to bind to Zone when user submits code to your API (which class stores executable code), not later, even if the . The user will expect
\item Be aware that if your API methods rely on other API method (overloaded method to provide default argument), you should not bind submitted code twice.
\item If your framework rely only on another executor (i.e. does not store any form of executable code), you can simply bind any code submitted to that executor, or better, use a zone-instrumented version of that executor.
\item Think Zoned. As long as you don't make an asynchronous execution, context is automatically preserved (same thread)
\item Be aware of race conditions. Zones are shared among concurrent parts of programs by nature + crossing hooks \emph{will} execute in foreign Zone : An object modified by a crossing hook is not Thread safe, even it is in a thread confined zone.
\end{itemize}